---
title: "Exploratory Data Analysis"
author: "Juan Orduz"
output:
  html_document:
    df_print: paged
---

This exploratory analysis could server as a first step towards a reporting template which might be used to track and visualize the status of the past and current students. Potentially, this couls also be displayed in a [Shiny Dashboard](https://rstudio.github.io/shinydashboard/). 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)
```

# Prepare Notebook

```{r}
library(glue)
library(igraph)
library(magrittr)
library(readxl)
library(tidytext)
library(tidyverse)
library(widyr)
library(wordcloud)

MAIN_DIR <- here::here()
```

# Read Data

```{r}
# Data Set 1
haupttabelle_raw_df <- read_csv(file = glue("{MAIN_DIR}/data/Haupttabelle_ueberarbeitet_20180424 - Tabelle1.csv"))

# Data Set 2
sheets <- excel_sheets(path = glue("{MAIN_DIR}/data/Survey Responses SOG.xlsx"))

survey_responses_raw <- sheets %>%
  keep(str_detect(string = ., pattern = "r")) %>% 
  map(.f = function(sheet) {
    read_excel(path = glue("{MAIN_DIR}/data/Survey Responses SOG.xlsx"), sheet = sheet)
  }
)
```

# Data Cleaning

```{r}
haupttabelle_df <- haupttabelle_raw_df %>% 
  mutate(success = case_when(
    Abschlussnote %in% c("D", "S") ~ "yes", # D and S are passed.
    Abschlussnote == "failed" ~ "no", 
    TRUE ~ NA_character_
  )) %>% 
  mutate_if(.predicate = is.character, .funs = str_to_lower) %>% 
  mutate(Abschlussnote = as_factor(Abschlussnote, level = c("d", "s", "failed", "ausstehend")))
```

```{r}
survey_responses_df <- survey_responses_raw %>% 
  map_df(.f = 
    ~ gather(., key = Frage, value = Answer, ... = - `Personen-ID`, - Jahr) %>% 
      unite(col = Jahr_Frage, ... = Jahr, Frage, sep = "_")
)

survey_responses_df %<>% mutate_if(.predicate = is.character, .funs = str_to_lower) 
```

# Combine Data 

```{r}
# TODO: Verify Unique_ID convention. 
all_data_df <- full_join(x = haupttabelle_df, y = survey_responses_df, by = c("Unique_ID" = "Personen-ID"))
```

# Descriptive Statistics

## Haupttabelle

```{r}
# Get number of unique students. 
haupttabelle_df %>% pull(Unique_ID) %>% unique() %>% length()
```

```{r, fig.align="center"}
# Plot distribution of studies duration for successful students.
haupttabelle_df %>% 
  filter(success == "yes") %>% 
  mutate(Studienabschlussjahr = as.numeric(Studienabschlussjahr)) %>% 
  mutate(studies_length = Studienabschlussjahr - Studienbeginn) %>% 
  ggplot() +
  geom_histogram(mapping = aes(studies_length), fill = "dark blue") +
  labs(title = "Successful Studies Duration Distribution")
  
```

```{r, fig.align="center"}
haupttabelle_df %>% 
  filter(Geburtsjahr > 1950) %>% # There are some outliers (e.g. 1905).
  ggplot(mapping = aes(x = Geburtsjahr, fill = Geschlecht)) +
  geom_density() +
  scale_fill_manual(values = c("dark blue", "dark red")) +
  facet_grid(cols = vars(Geschlecht)) +
  labs(title = "Birthdate Distribution per Gender", y = "")
```

```{r, fig.align="center"}
# Plot time development of new students per gender. 
haupttabelle_df %>% 
  group_by(Geschlecht, Studienbeginn) %>% 
  count() %>% 
  ggplot() +
  geom_line(mapping = aes(x = Studienbeginn, y = n, color = Geschlecht), alpha = 0.5) +
  geom_point(mapping = aes(x = Studienbeginn, y = n, color = Geschlecht)) +
  labs(title = "Studienbeginn Timeline") +
  scale_color_manual(values = c("dark blue", "dark red")) + 
  scale_x_discrete(limits = c(2009:2017))
  
```

```{r, fig.align="center"}
# Plot grade distribution of students per gender. 
#   - D and S mean pass (D is the best grade).
haupttabelle_df %>% 
  count(Geschlecht, Abschlussnote) %>% 
  ggplot() +
  geom_bar(mapping = aes(x = Abschlussnote, y = n, fill = Geschlecht), 
           stat = "identity") +
  labs(title = "Grade Distribution by Gender") +
  scale_fill_manual(values = c("dark blue", "dark red")) +
  scale_x_discrete(limits = c("d", "s", "failed", "ausstehend"))
```

```{r, fig.align="center"}
# Plot grade share of students per gender. 
haupttabelle_df %>% 
  count(Geschlecht, Abschlussnote) %>% 
  ggplot() +
  geom_bar(mapping = aes(x = Abschlussnote, y = n, fill = Geschlecht), 
           stat = "identity", 
           position = "fill") +
  labs(title = "Grade Share by Gender") +
  scale_fill_manual(values = c("dark blue", "dark red")) +
  scale_x_discrete(limits = c("d", "s", "failed", "ausstehend"))
```

```{r, fig.align="center"}
haupttabelle_df %>% 
  count(Abschlussnote, Geschlecht) %>% 
  spread(key = Geschlecht, value = n, fill = 0) %>% 
  rowwise() %>% 
  mutate(
    p = prop.test(x = weiblich, (weiblich + männlich))$estimate,
    ci_lwr = prop.test(x = weiblich, (weiblich + männlich))$conf.int[1],
    ci_upr = prop.test(x = weiblich, (weiblich + männlich))$conf.int[2]
  ) %>% 
  gather(key = Geschlecht, ... = c("weiblich", "männlich")) %>% 
  ggplot(mapping = aes(x = Abschlussnote, y = p)) +
  geom_point(color = "dark red") +
  geom_errorbar(mapping = aes(ymin = ci_lwr, ymax = ci_upr), 
                color = "dark red", 
                alpha = 0.5) +
  geom_hline(yintercept = 0.5, alpha = 0.5) +
  labs(title = "Women Grades Share", y = "Share") +
  scale_x_discrete(limits = c("d", "s", "failed", "ausstehend")) +
  ylim(c(0, 1))
```



```{r, fig.align="center"}
# Plot of the sucess ratio per gender with 95% confidence interval. 
haupttabelle_df %>% 
  # We filter for students which should have studied.
  filter(Studienbeginn < 2016) %>% 
  filter(!is.na(success)) %>% 
  count(Studienbeginn, Geschlecht, success) %>% 
  spread(key = success, value = n, fill = 0) %>% 
  rowwise() %>% 
  mutate(
    p = prop.test(x = yes, (yes + no))$estimate,
    ci_lwr = prop.test(x = yes, (yes + no))$conf.int[1],
    ci_upr = prop.test(x = yes, (yes + no))$conf.int[2]
  ) %>% 
  gather(key = success, ... = c("yes", "no")) %>% 
  ggplot(mapping = aes(x = Studienbeginn, y = p, color = Geschlecht)) +
  geom_point() +
  geom_errorbar(mapping = aes(ymin = ci_lwr, ymax = ci_upr), alpha = 0.5) +
  geom_line(alpha = 1) +
  scale_color_manual(values = c("dark blue", "dark red")) +
  facet_grid(rows = vars(Geschlecht)) +
  labs(title = "Success Ratio per Starting Year (Studienbeginn)", 
       y = "Success Rate")
```

```{r, fig.align="center"}
# Plot count of sucess as a function of birth date. 
haupttabelle_df %>% 
  filter(Geburtsjahr > 1960) %>% 
  filter(! is.na(success)) %>% 
  count(Geburtsjahr, success) %>% 
  ggplot() +
  geom_bar(mapping = aes(x = Geburtsjahr, y = n, fill = success), 
           stat = "identity", 
           position = "dodge") +
  labs(title = "Success Count") +
  scale_fill_manual(values = c("dark red", "dark green"))
```

```{r, fig.align="center"}
# Plot count of sucess as a function of birth date. 
haupttabelle_df %>% 
  filter(Geburtsjahr > 1960) %>% 
  filter(! is.na(success)) %>% 
  count(Geburtsjahr, Geschlecht, success) %>% 
  ggplot() +
  geom_bar(mapping = aes(x = Geburtsjahr, y = n, fill = success), 
           stat = "identity", 
           position = "dodge") +
  labs(title = "Success Count per Gernder") +
  scale_fill_manual(values = c("dark red", "dark green")) +
  facet_grid(rows = vars(Geschlecht))
```

```{r, fig.align="center"}
# Plot top words on the project categories. 
haupttabelle_df %>% 
  select(Projektkategorie_vereinfacht) %>% 
  unnest_tokens(output = word, input = Projektkategorie_vereinfacht) %>% 
  filter(! is.na(word)) %>% 
  count(word) %>% 
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n)) +
  geom_col(fill = "dark blue") +
  xlab(NULL) +
  coord_flip() +
  labs(title = "Projektkategorie - Word Count")
```


# Deep Dive

## Text Analysis: Remarks on Non-Successful Students

### Word Count

We want to undersand the remarks added to the data set associated to the students which were not successful. We use basic word statistics to explore this feature. 

```{r}
anmerkungen_fail_raw_df <- haupttabelle_df %>% 
  # filter for non-successful students. 
  filter(success == "no") %>% 
  select(Anmerkungen) %>% 
  rowid_to_column()
```

```{r}
# Get german stopwords. 
stopwords_df <- tibble(word = stopwords::stopwords(language = "de"))
# We want to keep some stop words. 
stopwords_to_keep <- c("nicht", "nichts")
stopwords_df %<>% filter(!(word %in% stopwords_to_keep))

# Get words. 
words_df <- anmerkungen_fail_raw_df %>% 
  unnest_tokens(output = word, input = Anmerkungen) %>% 
  anti_join(y = stopwords_df, by = "word")  %>% 
  filter(is.na(as.numeric(word))) %>% 
  rowwise() %>% 
  mutate(word_length = str_length(word)) %>% 
  filter(word_length > 1) %>% ungroup()

# Get words count. 
word_count <- words_df %>% 
  count(word) %>% 
  arrange(- n)
```

We begin by visualizing a wordcloud.

```{r, fig.width=8, fig.align="center"}
wordcloud(words = word_count$word, 
          freq = word_count$n, 
          min.freq = 1, 
          colors = brewer.pal(8, "Dark2"))
```


```{r, fig.align="center"}
# We now plot the count of the top words. 
word_count %>% 
  filter(n > 4) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n)) +
  geom_col(fill = "dark blue") +
  xlab(NULL) +
  coord_flip() +
  labs(title = "Anmerkungen -  Non Success - Word Count")
```

### Pairwise Word Count

We now count the words the pairwise word appearance within a same Anmerkung. We encode this into a weighted network (graph) to try to find topics. 

```{r, fig.width=50, fig.height=50, fig.align="center"}
# We save the network plot as a pdf file.
pdf("annmerkungen_non_success_word_network.pdf")

# Remove manually noisy words. 
# E.g. The sentence "(siehe Email CADEP 04.03.15)" appears often on the this column.
words_to_rm <- c("04.03.15", "mail", "email", "cadep", "siehe")

# Compute pairwise word count. 
pair_wise_count_df <- words_df %>% 
  filter(!(word %in% words_to_rm)) %>% 
  pairwise_count(item = word, feature = rowid) %>% 
  filter(item1 < item2) %>% arrange(item1, item2) %>% 
  arrange(- n)

# Set a threshold for visualization.
threshold <- 1
# Define network object. 
network <-  pair_wise_count_df %>%
  filter(n > threshold) %>%
  graph_from_data_frame(directed = FALSE)

# Store the degree.
V(network)$degree <- degree(graph = network)

# Compute the weight shares.
E(network)$width <- E(network)$n/max(E(network)$n)
 
# Plot the network.
plot(network, 
     vertex.color = 'lightblue',
     # Scale node size by degree.
     vertex.size = V(network)$degree / 15,
     vertex.label.color = 'black', 
     vertex.label.cex = 0.3, 
     vertex.label.dist = 0.5,
     edge.color = 'gray', 
     # Set edge width proportional to the weight relative value.
     edge.width = 1*E(network)$width,
     main = "Anmerkungen -  Non Success - Paiwise Word Count", 
     sub = glue('Weight Threshold: {threshold}'), 
     alpha = 80)

dev.off()
```

```{r, fig.width=10, fig.height=8, fig.align="center"}
plot(network, 
     vertex.color = 'lightblue',
     # Scale node size by degree.
     vertex.size = V(network)$degree / 15,
     vertex.label.color = 'black', 
     vertex.label.cex = 0.3, 
     vertex.label.dist = 0.5,
     edge.color = 'gray', 
     # Set edge width proportional to the weight relative value.
     edge.width = 1*E(network)$width,
     main = "Anmerkungen -  Non Success - Paiwise Word Count", 
     sub = glue('Weight Threshold: {threshold}'), 
     alpha = 50)
```


## Text Analysis: Personal Information

We want to see if there is any difference between the successful vs non-successful students personal information, which is encoded as free text. 

```{r}
# Let un compute the share of students with personal information available. 
haupttabelle_df %>% filter(!is.na(pesoenliche_Information)) %>% nrow() / nrow(haupttabelle_df)
```


```{r, fig.align="center"}
pesoenliche_info_raw_df <- haupttabelle_df %>% 
  select(success, pesoenliche_Information) %>% 
  filter(! is.na(success)) %>% 
  rowid_to_column()

# Plot top words per success tag. 
pesoenliche_info_raw_df %>% 
  unnest_tokens(output = word, input = pesoenliche_Information) %>% 
  filter(!is.na(word)) %>% 
  anti_join(y = stopwords_df, by = "word") %>% 
  count(success, word) %>% 
  mutate(word = reorder(word, n)) %>%
  filter(n > 1) %>% 
  ggplot(aes(word, n)) +
  geom_col(fill = "dark blue") +
  xlab(NULL) +
  coord_flip() +
  labs(title = "Personal Info by Success Tag - Word Count") +
  facet_grid(cols = vars(success))
```

## Text Analysis: Projektarbeitstitel

```{r, fig.width=8, fig.align="center"}
words_df <- haupttabelle_df %>% 
  select(Projektarbeitstitel) %>% 
  rowid_to_column() %>% 
  unnest_tokens(output = word, input = Projektarbeitstitel) %>% 
  anti_join(y = stopwords_df, by = "word")  %>% 
  filter(is.na(as.numeric(word))) %>% 
  rowwise() %>% 
  mutate(word_length = str_length(word)) %>% 
  filter(word_length > 1) %>% 
  ungroup()

# Get words count. 
word_count <- words_df %>% 
  count(word) %>% 
  arrange(- n)

wordcloud(words = word_count$word, 
          freq = word_count$n, 
          min.freq = 3, 
          colors = brewer.pal(8, "Dark2"))
```


```{r, fig.width=50, fig.height=50, fig.align="center"}
# We run a similar network analysis to create a visualization about the project titles. 

words_to_rm <- c("dass", "ab", "möchte", "geht", "sollen", "de", "mehr", "somit", "dabei", "darauf", "sowie")

# Compute pairwise word count. 
pair_wise_count_df <- words_df %>% 
  filter(!(word %in% words_to_rm)) %>% 
  pairwise_count(item = word, feature = rowid) %>% 
  filter(item1 < item2) %>% arrange(item1, item2) %>% 
  arrange(- n)

# Set a threshold for visualization.
threshold <- 2
# Define network object. 
network <-  pair_wise_count_df %>%
  filter(n > threshold) %>%
  graph_from_data_frame(directed = FALSE)

# Store the degree.
V(network)$degree <- degree(graph = network)

# Compute the weight shares.
E(network)$width <- E(network)$n/max(E(network)$n)
```

```{r, fig.width=50, fig.height=50, fig.align="center", eval=FALSE}
# We save the network plot as a pdf file.
pdf("projektarbeitstitel_word_network.pdf")
# Plot the network.
plot(network, 
     vertex.color = rgb(0,0,1, alpha = 0.2),
     # Scale node size by degree.
     vertex.size = 1.0*V(network)$degree,
     vertex.label.color = 'black', 
     vertex.label.cex = 0.3, 
     vertex.label.dist = 0.0,
     edge.color = 'gray', 
     # Set edge width proportional to the weight relative value.
     edge.width = 2*E(network)$width,
     main = "Projektarbeitstitel - Paiwise Word Count", 
     #sub = glue('Weight Threshold: {threshold}'), 
     alpha = 80)

dev.off()
```

```{r, fig.width=10, fig.height=8, fig.align="center"}
# Plot the network.
plot(network, 
     vertex.color = rgb(0,0,1, alpha = 0.2),
     # Scale node size by degree.
     vertex.size = 1.0*V(network)$degree,
     vertex.label.color = 'black', 
     vertex.label.cex = 0.3, 
     vertex.label.dist = 0.0,
     edge.color = 'gray', 
     # Set edge width proportional to the weight relative value.
     edge.width = 2*E(network)$width,
     main = "Projektarbeitstitel - Paiwise Word Count", 
     #sub = glue('Weight Threshold: {threshold}'), 
     alpha = 80)
```

